---
title: "Adding_2024_2024"
author: "Lana Bolin"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
---

# Parking downhill

* I need to do for spring data what I did for fall. See the "Make a mask" section in top chunk, and "### NEW ###" merging step. 
* I need to fix the fact that quad 4 was monitored some years. I gotta make rows of NAs, and figure out how to merge those.

```{r, message=F, warning=F}
rm(list = ls())

libraries <- c("tidyverse", "magrittr", "lme4", "car", "emmeans", "glmmTMB", "DHARMa", "multcomp", "ggeffects", "nlme", "MuMIn", "visreg", "ggsignif", "sjPlot", "sjmisc", "reshape2", "googlesheets4", "purrr", "stringr")
invisible(lapply(libraries, require, character.only = T))

setwd("~/GitHub/2025_lab_stomp")

# Load 2024/2025 STOMP data (has both sites)
new_data_long <- read.csv("~/GitHub/2025_lab_stomp/data/biocrust_2024_2025.csv", stringsAsFactors = T)




### Load corrected fall data (grass and creosote)
old_data_fall <- read.csv("~/GitHub/2025_lab_stomp/data/corrected.fall.data.wide.sp.csv")
old_data_spring <- read.csv("~/GitHub/2025_lab_stomp/data/corrected.spring.data.wide.sp.csv")

### ---------- BUILD OLD UNMONITORED MASK (PROPERLY STANDARDIZED) ----------

build_old_mask <- function(old_wide_df, season_label) {

  id_cols <- names(old_wide_df)[1:7]
  sp_cols <- names(old_wide_df)[8:ncol(old_wide_df)]

  old_wide_df %>%
    mutate(
      unmonitored = if_all(all_of(sp_cols), is.na)
    ) %>%
    filter(unmonitored) %>%
    dplyr::select(all_of(id_cols)) %>%
    distinct() %>%
    # ---- Apply SAME standardization logic as std_long() ----
    mutate(
      treatment = dplyr::recode(treatment,
                          "C" = "Ambient",
                          "STOMP" = "Degraded"),
site = dplyr::recode(site,
                     "crust_grass" = "Grassland",
                     "crust_creosote" = "Shrubland"),
      season = season_label,

      treatment = factor(treatment, levels = c("Ambient","Degraded")),
      season    = factor(season, levels = c("Spring","Fall")),
      site      = factor(site, levels = c("Grassland","Shrubland")),

      year    = as.integer(year),
      plot    = as.integer(plot),
      quad    = as.integer(quad),
      quad_id = as.factor(quad_id)
    )
}

# Build masks
old_unmonitored_keys_fall   <- build_old_mask(old_data_fall,   "Fall")
old_unmonitored_keys_spring <- build_old_mask(old_data_spring, "Spring")


```

First, fix quad_id levels in the new data so they match new data. So "crust_creosote-NA-2-1-NA-NA-C" would become "crust_creosote_C_2_1".
```{r}
#library(stringr)
#library(purrr)

new_data_long <- new_data_long %>%
  mutate(
    quadID = quadID %>%
      str_split("-") %>%
      map_chr(function(x) {

        # remove NA + "NA"
        x <- x[!is.na(x) & x != "NA"]

        # if nothing left, return NA
        if (length(x) == 0) return(NA_character_)

        # identify parts
        first_part  <- x[1]
        last_part   <- x[length(x)]
        middle_part <- x[2:(length(x)-1)]

        # keep only numeric middle parts
        middle_part <- middle_part[str_detect(middle_part, "^[0-9]+$")]

        # reorder
        new_vec <- c(first_part, last_part, middle_part)

        str_c(new_vec, collapse = "_")
      })
  )
```


Make old data long format. 
```{r}
### Spring
sp_cols <- names(old_data_spring)[8:ncol(old_data_spring)]

old_data_spring_long <- old_data_spring %>%
  tidyr::pivot_longer(
    cols = all_of(sp_cols),
    names_to = "kartez",
    values_to = "cover"
  ) %>%
  dplyr::mutate(cover = suppressWarnings(as.numeric(as.character(cover))))



### Fall
sp_cols <- names(old_data_fall)[8:ncol(old_data_fall)]

old_data_fall_long <- old_data_fall %>%
  tidyr::pivot_longer(
    cols = all_of(sp_cols),
    names_to = "kartez",
    values_to = "cover"
  ) %>%
  dplyr::mutate(cover = suppressWarnings(as.numeric(as.character(cover))))
```


Fix columns, fix sporo and chama and astra, remove "EMPTY"
```{r}
# Re-name sites to match our old data
new_data_long <- new_data_long %>%
  dplyr::mutate(
    site = dplyr::recode(
      site,
      "crust_creosote" = "Shrubland",
      "crust_grass"    = "Grassland"
    ),
    treatment = dplyr::recode(
      treatment,
      "C"     = "Ambient",
      "STOMP" = "Degraded"
    ),
    season = dplyr::recode(
      season,
      "2" = "Spring",
      "3" = "Fall"
    )
  )


new_data_long %<>%
  mutate(
    quadID = as.character(quadID)  # handle factor/character differences up front
  ) %>%
  dplyr::rename(
    kartez  = species,
    quad_id = quadID
  ) %>%
  mutate(
    quad_id = as.character(quad_id),
    kartez  = as.character(kartez),
    cover   = as.numeric(cover),
    treatment = as.character(treatment),
    season    = as.character(season),
    site      = as.character(site)
  ) %>%
  dplyr::select(
    treatment, year, season, site, plot, quad, quad_id, kartez, cover
  )




### Fix SPORO and CHAMA, remove EMPTY

new_data_long %<>% filter(kartez != "EMPTY")

sporo <- new_data_long %>%
  filter(kartez %in% c("SPCO4", "SPCR", "SPFL2", "SPORO")) %>%
  group_by(year, season, site, treatment, plot, quad, quad_id) %>%
  dplyr::summarise(cover = sum(cover, na.rm = TRUE), .groups = "drop")
sporo$kartez <- "SPORO"
nrow(sporo)

chama <- new_data_long %>%
  filter(kartez %in% c("CHAL11","CHAMA15","CHCHC3","CHFE3","CHAL10","CHLA10","CHRE4","CHSE7","CHSES","CHAMA","CHST8","CHSE6","CHCH5")) %>%
  group_by(year, season, site, treatment, plot, quad, quad_id) %>%
  dplyr::summarise(cover = sum(cover, na.rm = TRUE), .groups = "drop")
chama$kartez <- "CHAMA"
nrow(chama)

new_data_long %<>% dplyr::select(year, season, site, treatment, plot, quad, quad_id, cover, kartez)
nrow(new_data_long)


# Change ASNU4 and ASFE2 to ASTRA
astra <- new_data_long %>%
  filter(kartez %in% c("ASNU4", "ASFE2")) %>%
  group_by(year, season, site, treatment, plot, quad, quad_id) %>%
  dplyr::summarise(cover = sum(cover, na.rm = TRUE), .groups = "drop")
astra$kartez <- "ASTRA"

# Remove species that have been combined into SPORO and CHAMA and ASTRA
new_data_long %<>%
  filter(!kartez %in% c("SPCO4", "SPCR", "SPFL2", "SPORO", "CHAL11", "CHAMA15", "CHCHC3", "CHFE3", "CHAL10", "CHLA10", "CHRE4", "CHSE7", "CHSES", "CHAMA", "CHST8", "CHSE6", "CHCH5", "ASNU4", "ASFE2"))
nrow(new_data_long)

# Good
new_data_long %>% filter(kartez %in% c("SPCO4", "SPCR", "SPFL2",  "CHAL11", "CHAMA15", "CHCHC3", "CHFE3", "CHAL10", "CHLA10", "CHRE4", "CHSE7", "CHSES", "CHST8", "CHSE6", "CHCH5", "ASNU4", "ASFE2"))



# Merge new SPORO and CHAMA data with all the other species
new_data_long <- dplyr::bind_rows(new_data_long, sporo, chama, astra)

nrow(new_data_long)

# good
new_data_long %>% filter(kartez %in% c("SPCO4", "SPCR", "SPFL2",  "CHAL11", "CHAMA15", "CHCHC3", "CHFE3", "CHAL10", "CHLA10", "CHRE4", "CHSE7", "CHSES", "CHST8", "CHSE6", "CHCH5", "ASNU4", "ASFE2"))
```


De-duplicate new data
```{r}
# --- helper: collapse duplicates (sum) but keep "all NA" as NA ---
dedup_sum_keepNA <- function(df) {
  df %>%
    dplyr::group_by(treatment, year, season, site, plot, quad, quad_id, kartez) %>%
    dplyr::summarise(
      cover = if (all(is.na(cover))) NA_real_ else sum(cover, na.rm = TRUE),
      .groups = "drop"
    )
}

# --- De-duplicate within each season (add up all individuals of the same species in a quad x year) ---
new_data_long_nodup <- dedup_sum_keepNA(new_data_long)
#old_spring_long_nodup <- dedup_sum_keepNA(old_data_spring_long)
#old_fall_long_nodup <- dedup_sum_keepNA(old_data_fall_long)

nrow(new_data_long)
nrow(new_data_long_nodup)
```

Split new data by season, pivot wide
```{r}
# --- 1) Split into Spring / Fall long ---
new_spring_long_nodup <- new_data_long_nodup %>%
  dplyr::filter(season == "Spring")

new_fall_long_nodup <- new_data_long_nodup %>%
  dplyr::filter(season == "Fall")


# --- 3) Pivot each season to wide, fill with 0s ---
to_wide_by_kartez <- function(df_long) {
  df_long %>%
    tidyr::pivot_wider(
      names_from  = kartez,
      values_from = cover,
      values_fill = 0
    )
}

new_spring_wide <- to_wide_by_kartez(new_spring_long_nodup)
new_fall_wide   <- to_wide_by_kartez(new_fall_long_nodup)
old_spring_wide <- to_wide_by_kartez(old_data_spring_long)
old_fall_wide <- to_wide_by_kartez(old_data_fall_long)

# Check for no duplicates left (should be all 1s)
new_spring_long_nodup %>%
  dplyr::count(treatment, year, season, site, plot, quad, quad_id, kartez) %>%
  dplyr::summarise(max_n = max(n))

new_fall_long_nodup %>%
  dplyr::count(treatment, year, season, site, plot, quad, quad_id, kartez) %>%
  dplyr::summarise(max_n = max(n))

old_data_spring_long %>%
  dplyr::count(treatment, year, season, site, plot, quad, quad_id, kartez) %>%
  dplyr::summarise(max_n = max(n))

old_data_fall_long %>%
  dplyr::count(treatment, year, season, site, plot, quad, quad_id, kartez) %>%
  dplyr::summarise(max_n = max(n))

```


Merge old and new (for each season) - FIX
```{r}
std_long <- function(df) {
  df %>%
    mutate(
      treatment = case_when(
        treatment %in% c("C","STOMP") ~ dplyr::recode(treatment, "C"="Ambient","STOMP"="Degraded"),
        TRUE ~ as.character(treatment)
      ),
      season = case_when(
        season %in% c("2","3") ~ dplyr::recode(as.character(season), "2"="Spring","3"="Fall"),
        TRUE ~ as.character(season)
      ),
      site = case_when(
        site %in% c("crust_grass","crust_creosote") ~ dplyr::recode(site, "crust_grass"="Grassland","crust_creosote"="Shrubland"),
        TRUE ~ as.character(site)
      ),

      treatment = factor(treatment, levels = c("Ambient", "Degraded")),
      season    = factor(season, levels = c("Spring", "Fall")),
      site      = factor(site, levels = c("Grassland", "Shrubland")),

      year    = as.integer(year),
      plot    = as.integer(plot),
      quad    = as.integer(quad),
      quad_id = as.character(quad_id),
      kartez  = as.character(kartez),
      cover   = as.numeric(cover)
    ) %>%
    dplyr::select(treatment, year, season, site, plot, quad, quad_id, kartez, cover)
}

all_long <- dplyr::bind_rows(
  std_long(old_data_fall_long),
  std_long(old_data_spring_long),
  std_long(new_spring_long_nodup),
  std_long(new_fall_long_nodup)
)

summary(all_long)
```

Change SPHAE to SPHA
```{r}
### Before going wide, look at fall quads with SPHAE and see if there are previous Sphaeralcea species. Unless the new observation is very small cover = likely to be a seedling, or if Sphaeralcea was previously absent from the quadrat (= new species observed).

# Find quads where SPHAE is ever present
quad_ids_sphae <- all_long %>%
  filter(season == "Fall") %>%
  dplyr::filter(kartez == "SPHAE", !is.na(cover), cover > 0) %>%
  dplyr::distinct(quad_id) %>%
  dplyr::pull(quad_id)

# Pull the full time series for those quads (all species, all years)
sphae_ts_long <- all_long %>%
  dplyr::filter(quad_id %in% quad_ids_sphae) %>%
  dplyr::arrange(quad_id, year, kartez)

# Look at each quad's time series: For all, change "SPHAE" to "SPHA"
sphae_ts_long %>% filter(quad_id == quad_ids_sphae[1]) %>%
  filter(str_starts(kartez, "SPH"))

sphae_ts_long %>% filter(quad_id == quad_ids_sphae[2]) %>%
  filter(str_starts(kartez, "SPH"))

sphae_ts_long %>% filter(quad_id == quad_ids_sphae[3]) %>%
  filter(str_starts(kartez, "SPH"))

sphae_ts_long %>% filter(quad_id == quad_ids_sphae[4]) %>%
  filter(str_starts(kartez, "SPH"))

sphae_ts_long %>% filter(quad_id == quad_ids_sphae[5]) %>%
  filter(str_starts(kartez, "SPH"))

sphae_ts_long %>% filter(quad_id == quad_ids_sphae[6]) %>%
  filter(str_starts(kartez, "SPH"))

all_long %<>%
  dplyr::mutate(
    kartez = dplyr::recode(kartez, "SPHAE" = "SPHA")
  )
```

Separate by season
```{r}
all_spring_long <- all_long %>% filter(season == "Spring")
all_fall_long <- all_long %>% filter(season == "Fall")
```

Go wide. There are some rows that have some NAs here for some reason, but I'll fix that in the next section.
```{r}
# Fall
all_fall_wide <- all_fall_long %>%
  pivot_wider(
    names_from  = kartez,
    values_from = cover,
    values_fill = 0
  )

# Spring
all_spring_wide <- all_spring_long %>%
  pivot_wider(
    names_from  = kartez,
    values_from = cover,
    values_fill = 0
  )
```

Make sure I have 4 quads for each plot x year in 2024 and 2025. Then I'll change the unmonitored ones to 0.
```{r}
new_years <- 2024:2025

ensure_four_quads <- function(df, sp_cols) {

  df %>%
    group_by(site, treatment, plot, year) %>%
    tidyr::complete(
      quad = 1:4,
      fill = as.list(setNames(rep(0, length(sp_cols)), sp_cols))
    ) %>%
    ungroup() %>%
    mutate(
      season = first(season),  # keep season column coherent if present

      site_code = case_when(
        site == "Grassland" ~ "crust_grass",
        site == "Shrubland" ~ "crust_creosote",
        TRUE ~ as.character(site)
      ),
      trt_code = case_when(
        treatment == "Ambient"  ~ "C",
        treatment == "Degraded" ~ "STOMP",
        TRUE ~ as.character(treatment)
      ),
      quad_id = paste(site_code, trt_code, plot, quad, sep = "_")
    ) %>%
    dplyr::select(-site_code, -trt_code)
}

# Split old vs new years
spring_old <- all_spring_wide %>% filter(year < 2024)
spring_new <- all_spring_wide %>% filter(year %in% new_years)

fall_old <- all_fall_wide %>% filter(year < 2024)
fall_new <- all_fall_wide %>% filter(year %in% new_years)


# Only enforce 4 quads on new years
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")
sp_cols_fall <- setdiff(names(all_fall_wide), id_cols)
sp_cols_spring <- setdiff(names(all_spring_wide), id_cols)

fall_new <- ensure_four_quads(fall_new, sp_cols_fall)
spring_new <- ensure_four_quads(spring_new, sp_cols_spring)

# Recombine
all_spring_wide <- bind_rows(spring_old, spring_new) %>%
  arrange(site, treatment, plot, year, quad)


all_fall_wide <- bind_rows(fall_old, fall_new) %>%
  arrange(site, treatment, plot, year, quad)

```



Force unmonitored quads to NA. Use the mask I created for the old data. For the new data, plot 29 should have NAs for quads 2 and 3, while all other plots should have NAs for quads 1 and 4.
```{r}
old_years <- 2014:2023
new_years <- 2024:2025

### Spring
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")
sp_cols_spring <- setdiff(names(all_spring_wide), id_cols)

# 1) Attach old mask
all_spring_wide <- all_spring_wide %>%
  left_join(
    old_unmonitored_keys_spring %>% mutate(old_unmonitored = TRUE),
    by = id_cols
  ) %>%
  mutate(old_unmonitored = coalesce(old_unmonitored, FALSE))

# 2. Apply conditional masking
all_spring_wide <- all_spring_wide %>%
  mutate(
    new_unmonitored = case_when(
      year %in% new_years & plot == 29 & quad %in% c(2,3) ~ TRUE,
      year %in% new_years & plot != 29 & quad %in% c(2,4) ~ TRUE,
      TRUE ~ FALSE
    ),
    
    final_unmonitored = case_when(
      year %in% old_years ~ old_unmonitored,
      year %in% new_years ~ new_unmonitored,
      TRUE ~ FALSE
    )
  ) %>%
  mutate(
    across(all_of(sp_cols_spring),
           ~ if_else(final_unmonitored, NA_real_, .))
  ) %>%
  dplyr::select(-old_unmonitored, -new_unmonitored, -final_unmonitored)


### Fall
sp_cols_fall <- setdiff(names(all_fall_wide), id_cols)

all_fall_wide <- all_fall_wide %>%
  left_join(
    old_unmonitored_keys_fall %>% mutate(old_unmonitored = TRUE),
    by = id_cols
  ) %>%
  mutate(old_unmonitored = coalesce(old_unmonitored, FALSE)) %>%
  mutate(
    new_unmonitored = case_when(
      year %in% new_years & plot == 29 & quad %in% c(2,3) ~ TRUE,
      year %in% new_years & plot != 29 & quad %in% c(2,4) ~ TRUE,
      TRUE ~ FALSE
    ),
    
    final_unmonitored = case_when(
      year %in% old_years ~ old_unmonitored,
      year %in% new_years ~ new_unmonitored,
      TRUE ~ FALSE
    )
  ) %>%
  mutate(
    across(all_of(sp_cols_fall),
           ~ if_else(final_unmonitored, NA_real_, .))
  ) %>%
  dplyr::select(-old_unmonitored, -new_unmonitored, -final_unmonitored)


### Sanity check - SPRING

# Rows should be all NA or no NA (no partial NA)
spring_check <- all_spring_wide %>%
  mutate(
    n_na = rowSums(is.na(across(all_of(sp_cols_spring)))),
    n_sp = length(sp_cols_spring),
    type = case_when(
      n_na == 0 ~ "no_NA",
      n_na == n_sp ~ "all_NA",
      TRUE ~ "partial_NA"
    )
  ) %>%
  count(type)

# Should see "all_NA" and "no_NA". No "partial_NA".
spring_check

# Check for 2 quads each year
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")
sp_cols_spring <- setdiff(names(all_spring_wide), id_cols)

spring_quad_check <- all_spring_wide %>%
  mutate(
    monitored = !if_all(all_of(sp_cols_spring), is.na)
  ) %>%
  group_by(site, treatment, plot, year) %>%
  summarise(
    n_monitored = sum(monitored),
    .groups = "drop"
  )

# Want n_monitored = 2 for all years
spring_quad_check %>%
  filter(year %in% 2014:2023) %>%
  count(n_monitored)

spring_quad_check %>%
  filter(year %in% 2024:2025) %>%
  count(n_monitored)

# Confirm no partial NAs left - want only TRUE
all_spring_wide %>%
  mutate(
    n_na = rowSums(is.na(across(all_of(sp_cols_spring)))),
    n_sp = length(sp_cols_spring)
  ) %>%
  count(n_na == 0 | n_na == n_sp)


### Sanity check - Fall

# Rows should be all NA or no NA (no partial NA)
fall_check <- all_fall_wide %>%
  mutate(
    n_na = rowSums(is.na(across(all_of(sp_cols_fall)))),
    n_sp = length(sp_cols_fall),
    type = case_when(
      n_na == 0 ~ "no_NA",
      n_na == n_sp ~ "all_NA",
      TRUE ~ "partial_NA"
    )
  ) %>%
  count(type)

# Should see "all_NA" and "no_NA". No "partial_NA".
fall_check

# Check for 2 quads each year
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")
sp_cols_fall <- setdiff(names(all_fall_wide), id_cols)

fall_quad_check <- all_fall_wide %>%
  mutate(
    monitored = !if_all(all_of(sp_cols_fall), is.na)
  ) %>%
  group_by(site, treatment, plot, year) %>%
  summarise(
    n_monitored = sum(monitored),
    .groups = "drop"
  )

# Want n_monitored = 2 for all years
fall_quad_check %>%
  filter(year %in% 2014:2023) %>%
  count(n_monitored)

fall_quad_check %>%
  filter(year %in% 2024:2025) %>%
  count(n_monitored)

# Confirm no partial NAs left - want only TRUE
all_fall_wide %>%
  mutate(
    n_na = rowSums(is.na(across(all_of(sp_cols_fall)))),
    n_sp = length(sp_cols_fall)
  ) %>%
  count(n_na == 0 | n_na == n_sp)
```

Check which quads were monitored each year in plot 29 
```{r}
# Spring
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")
sp_cols_spring <- setdiff(names(all_spring_wide), id_cols)

plot29_spring_monitoring <- all_spring_wide %>%
  filter(plot == 29) %>%
  mutate(
    monitored = !if_all(all_of(sp_cols_spring), is.na)
  ) %>%
  dplyr::select(site, treatment, year, quad, quad_id, monitored) %>%
  arrange(site, treatment, year, quad)

plot29_spring_monitoring


# Fall
sp_cols_fall <- setdiff(names(all_fall_wide), id_cols)

plot29_fall_monitoring <- all_fall_wide %>%
  filter(plot == 29) %>%
  mutate(
    monitored = !if_all(all_of(sp_cols_fall), is.na)
  ) %>%
  dplyr::select(site, treatment, year, quad, quad_id, monitored) %>%
  arrange(site, treatment, year, quad)

plot29_fall_monitoring
```

Fix season and quad_id columns for the fake rows I created
```{r}
# Spring
all_spring_wide <- all_spring_wide %>%
  mutate(season = factor("Spring", levels = c("Spring","Fall")))

all_spring_wide <- all_spring_wide %>%
  mutate(
    # convert site back to original code
    site_code = case_when(
      site == "Grassland" ~ "crust_grass",
      site == "Shrubland" ~ "crust_creosote"
    ),
    
    # convert treatment back to original code
    trt_code = case_when(
      treatment == "Ambient" ~ "C",
      treatment == "Degraded" ~ "STOMP"
    ),
    
    quad_id = paste(site_code, trt_code, plot, quad, sep = "_")
  ) %>%
  dplyr::select(-site_code, -trt_code)

all_fall_wide %>% filter(is.na(quad_id))


# Fall
all_fall_wide <- all_fall_wide %>%
  mutate(season = factor("Fall", levels = c("Spring","Fall")))

all_fall_wide <- all_fall_wide %>%
  mutate(
    # convert site back to original code
    site_code = case_when(
      site == "Grassland" ~ "crust_grass",
      site == "Shrubland" ~ "crust_creosote"
    ),
    
    # convert treatment back to original code
    trt_code = case_when(
      treatment == "Ambient" ~ "C",
      treatment == "Degraded" ~ "STOMP"
    ),
    
    quad_id = paste(site_code, trt_code, plot, quad, sep = "_")
  ) %>%
  dplyr::select(-site_code, -trt_code)

all_fall_wide %>% filter(is.na(quad_id))
```


Look for rows that all are 0
```{r}
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")
sp_cols_spring <- setdiff(names(all_spring_wide), id_cols)
sp_cols_fall   <- setdiff(names(all_fall_wide),   id_cols)

spring_all_zero <- all_spring_wide %>%
  mutate(
    all_zero = if_all(all_of(sp_cols_spring), ~ . == 0)
  ) %>%
  filter(all_zero) %>%
  dplyr::select(all_of(id_cols)) %>%
  arrange(site, treatment, plot, year, quad)

spring_all_zero
nrow(spring_all_zero)

fall_all_zero <- all_fall_wide %>%
  mutate(
    all_zero = if_all(all_of(sp_cols_fall), ~ . == 0)
  ) %>%
  filter(all_zero) %>%
  dplyr::select(all_of(id_cols)) %>%
  arrange(site, treatment, plot, year, quad)

fall_all_zero
nrow(fall_all_zero)
```

```{r}
check_plot <- function(df, id_cols, plot_num) {
  sp_cols <- setdiff(names(df), id_cols)
  df %>%
    dplyr::filter(plot == plot_num, year %in% 2024:2025) %>%
    dplyr::mutate(monitored = !dplyr::if_all(dplyr::all_of(sp_cols), is.na)) %>%
    dplyr::select(site, treatment, season, year, plot, quad, quad_id, monitored) %>%
    dplyr::arrange(site, treatment, year, quad)
}

check_plot(all_spring_wide, id_cols, 29)
check_plot(all_spring_wide, id_cols, 25)
```
I convinced myself these are real ecological zeros, or empty quads.


Fix species names from Mel
```{r}
rename_species_wide <- function(df, id_cols) {
  
  sp_map <- c(
    "CHAMA"  = "CHAMA15",
    "DRCUC"  = "DRCU",
    "ECFEF3" = "ECFE",
    "GASUN"  = "OESU4",
    "HYFIC"  = "HYFI",
    "LAOCO"  = "LAOC3",
    "MAPIP"  = "MAPI",
    "MUAR"   = "MUAR2",
    "PSTAT"  = "PSTA",
    "SEFLF"  = "SEFL3",
    "TRRA"   = "TRBE"
  )
  
  for (old in names(sp_map)) {
    
    new <- sp_map[[old]]
    
    if (old %in% names(df)) {
      
      # if target column already exists, sum them
      if (new %in% names(df)) {
        a <- df[[new]]
        b <- df[[old]]
        
        df[[new]] <- ifelse(
          is.na(a) & is.na(b),
          NA_real_,
          rowSums(cbind(a, b), na.rm = TRUE)
        )
      } else {
        names(df)[names(df) == old] <- new
      }
      
      # remove old column if still present
      df[[old]] <- NULL
    }
  }
  
  df
}

id_cols <- c("treatment","year","season","site","plot","quad","quad_id")

all_spring_wide <- rename_species_wide(all_spring_wide, id_cols)
all_fall_wide   <- rename_species_wide(all_fall_wide,   id_cols)

# Sanity check
intersect(names(all_spring_wide), c("CHAMA","DRCUC","ECFEF3","GASUN","HYFIC",
                                    "LAOCO","MAPIP","MUAR","PSTAT","SEFLF","TRRA"))
```

Drop rows that are all NAs before going long and exporting (so lab doesn't need to deal with the NA rows)
```{r}
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")

all_spring_wide_clean <- all_spring_wide %>%
  filter(!if_all(-all_of(id_cols), is.na))

all_fall_wide_clean <- all_fall_wide %>%
  filter(!if_all(-all_of(id_cols), is.na))


# Sanity check - number of rows dropped
nrow(all_spring_wide) - nrow(all_spring_wide_clean)
nrow(all_fall_wide)   - nrow(all_fall_wide_clean)

# Confirm no NA rows remain
all_spring_wide_clean %>%
  filter(if_all(-all_of(id_cols), is.na)) %>%
  nrow()

all_fall_wide_clean %>%
  filter(if_all(-all_of(id_cols), is.na)) %>%
  nrow()
```


Go to long form
```{r}
id_cols <- c("treatment","year","season","site","plot","quad","quad_id")

all_spring_long2 <- all_spring_wide_clean %>%
  tidyr::pivot_longer(
    cols = -all_of(id_cols),
    names_to = "kartez",
    values_to = "cover"
  ) %>%
  dplyr::arrange(site, treatment, plot, year, quad, kartez)

all_fall_long2 <- all_fall_wide_clean %>%
  tidyr::pivot_longer(
    cols = -all_of(id_cols),
    names_to = "kartez",
    values_to = "cover"
  ) %>%
  dplyr::arrange(site, treatment, plot, year, quad, kartez)
```




Write csv for long and wide format for each season
```{r}
#write.csv(fall_wide, "./data/fall_wide_sp_allyears.csv", row.names = FALSE)
#write.csv(spring_wide, "./data/spring_wide_sp_allyears.csv", row.names = FALSE)

write.csv(all_fall_long2, "./data/fall_long_allyears.csv", row.names = FALSE)
write.csv(all_spring_long2, "./data/spring_long_allyears.csv", row.names = FALSE)

write.csv(all_fall_wide_clean, "./data/fall_wide_sp_allyears.csv")
write.csv(all_spring_wide_clean, "./data/spring_wide_sp_allyears.csv")
```



```{r}
all_spring_wide_clean %>% filter(year == 2014, plot == 28) %>% distinct(quad)
```

Make supplemental table showing which quads were monitored when. These are the years x plots x seasons where quads other than 1 and 3 were monitored
```{r}
# Fall
quads_by_plot_year_fall <- all_fall_long2 %>%
  distinct(plot, year, quad) %>%
  group_by(plot, year) %>%
  summarise(
    quads_monitored = paste(sort(unique(quad)), collapse = ", "),
    n_quads = n_distinct(quad),
    .groups = "drop"
  )

quads_by_plot_year_fall %>% filter(quads_monitored != "1, 3")

# Spring
quads_by_plot_year_spring <- all_spring_long2 %>%
  distinct(plot, year, quad) %>%
  group_by(plot, year) %>%
  summarise(
    quads_monitored = paste(sort(unique(quad)), collapse = ", "),
    n_quads = n_distinct(quad),
    .groups = "drop"
  )

quads_by_plot_year_spring %>% filter(quads_monitored != "1, 3")
```











